---
title: "03 - Kernel (Density Plots)"
author: "Francisco E. Fonturbel and Javier Cuadra"
date: "6/26/2019"
output: 
  html_document:
    includes:
        after_body: footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Kernel Density Plots

## Kenrnel density plots are effective way to view distribution of variable (time). Can be considered as plots of smoothed histograms


```{r library, echo=FALSE}
library ("overlap")
library ("activity")
```

## load data (time)

```{r}

act_x<-read.table("./Data/act_x.txt", header= T)

act_y<-read.table("./Data/act_y.txt", header= T)


```

## transform data (time) to radian

```{r}

rad_x<-act_x$Time_2*2*pi
rad_y<-act_y$Time_2*2*pi

```

## fitted data y extraction for condition 



```{r}
fitact_a <-fitact(rad_x[act_x$Condition=="A"])
fitact_b <-fitact(rad_y[act_Y$Condition=="B"])
```

# Density Plot


```{r}
plot(fitact_a, yunit="dens", las=1, dline=list(col="lightblue"), tline=list(col="blue"), ylim=c(0,0.17))

```

```{r}
plot(fitact_b, yunit="dens", las=1, dline=list(col="pink"), tline=list(col="red"))
     
```


## plot two graphs in same plot

```{r}
plot(fitact_a, yunit="dens", las=1, dline=list(col="lightblue"), tline=list(col="blue"), ylim=c(0,0.17))
     plot(fitact_a, yunit="dens", add=TRUE, dline=list(col="pink"), tline=list(col="red"))
     legend("topright", inset = c(0,0), title = "Periodo", c("A", "B"), lty= c(1,3), col = c("lightblue", "pink"), bty = "n")

```





Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
