---
title: "03 - Kernel (Density Plots)"
author: "Francisco E. Fonturbel and Javier Cuadra"
date: "6/26/2019"
output: 
  html_document:
    includes:
        after_body: footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Kernel Density Plots


## Kenrnel density plots are effective way to view distribution of variable (time). Can be considered as plots of smoothed histograms

# Package 'activity'


```{r library, echo=FALSE}

library ("activity")
```

## load data (time)

```{r}

act_x<-read.table("./Data/act_x.txt", header= T)

act_y<-read.table("./Data/act_y.txt", header= T)


```

## transform data (time) to radian

```{r}

rad_x<-act_x$Time_2*2*pi
rad_y<-act_y$Time_2*2*pi

```

## fitted data y extraction for condition 



```{r}
fitact_a <-fitact(rad_x[act_x$Condition=="A"])
fitact_b <-fitact(rad_y[act_y$Condition=="A"])
```

## Density Plot


```{r}
plot(fitact_a, yunit="dens", las=1, dline=list(col="lightblue"), tline=list(col="blue"), main="Data X", ylim=c(0,0.17))

```

```{r}
plot(fitact_b, yunit="dens", las=1, dline=list(col="pink"), tline=list(col="red"), main = "Data Y")
     
```


## plot two graphs in same plot. Compare same condition (A)

```{r}
plot(fitact_a, yunit="dens", las=1, dline=list(col="lightblue"), tline=list(col="blue"), ylim=c(0,0.17))
     plot(fitact_b, yunit="dens", add=TRUE, dline=list(col="pink"), tline=list(col="red"))
     legend("topright", inset = c(0,0), title = "Data", c("X", "Y"), lty= c(1,1), col = c("lightblue", "pink"), bty = "n")

```


# Coefficient of overlap (Dhat5)

## Coefficient of is the area under a density curve.This index is purely descriptive.

## The function 'CompareCkern' in the package 'activity', is a test probability that two sets of data come from the same distribution. In addition, this test calculate overlap index Dhat5 for two fitted distributions (with function 'fitact')

## Compare data X and Y, for same condition  "A"
```{r}
compareCkern(fitact_a, fitact_b, reps = 100)
```






# Package Overlap

## extraction of data for condition 

```{r}

act_a <- rad_x[act_x$Condition=="A"]

act_b <- rad_y[act_y$Condition=="A"]

```

## Density Plot

```{r}
library(overlap)
densityPlot(act_a, xscale = 24, xcenter = c("noon", "midnight"), add = F, rug = T, extend = "lightgray", main = "Data X", xlab = "Time", ylab = "Density", bty = "L")


```

```{r}
densityPlot(act_b, xscale = 24, xcenter = c("noon", "midnight"), add = F, rug = T, extend = "lightgray", main = "Data Y", xlab = "Time", ylab = "Density", bty = "L")
```

## plot two graphs in same plot with 'Overlap', for same condition (A)

```{r}
overlapPlot(act_a, act_b, main = " ", linecol = c("black", "darkgreen"), lty = c(1,3), bty = "L", xlab = "Time", ylab = "Density")
legend("topright", inset = c(0,0), title = "Data", c("X", "Y"), lty= c(1,3), col = c("gray25", "seagreen"), bty = "n")

```




Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
